Solving Collective Spin Hamiltonians with Decoherence

### 2018-08-20 soc_squeezing/cumulant_expansions.tex
update to last comutator derivation

We now consider again squeezing the initial state $\ket{X}$, but this
time with the transverse field Hamiltonian
\begin{align}
  H_{\t{TVF}}
  \equiv \chi S_z^2- \Omega S_x.
\end{align}
To determine the coherent evolution of the operator
$S_\mu^m S_\nu^n S_z^n$, we now need to compute
\begin{align}
  \sp{H_{\t{TVF}}, S_\mu^m S_\nu^n S_z^n}_-
  = \chi \sp{S_z^2, S_\mu^m S_\nu^n S_z^n}_-
  - \Omega \sp{S_x, S_\mu^m S_\nu^n S_z^n}_-.
\end{align}
Here first commutator is identical to that for OAT, while for the
second we need to compute
\begin{align}
  \sp{S_\rho, S_\mu^m S_\nu^n S_z^n}_-
\end{align}
for $\rho\in\set{+1,-1}$.

@@

@@

  \label{eq:E_mn_1_first}

@@

  = n^{\underline{p}}

@@

  D_{\mu,m-p,n-p}^{(k_p)} \prod_{j=1}^p \p{2k_j+j-n+2\mu S_z},
with $n^{\underline{p}}\equiv n\p{n-1}\cdots\p{n-p+1}$.  Observing
that we always have $k_1\le k_2\le\cdots\le k_p$, we can rearrange the
sums to get that for $p>1$
  = n^{\underline{p}}
  \sum_{\ell=0}^{m-p-1} D_{\mu,m-p,n-p}^{(\ell)} \p{2\ell+p-n+2\mu S_z}
  \sum_{\substack{\v k\in\Z_{\ell+1}^{p-1}\\k_1\le k_2\le\cdots\le k_{p-1}}}
  \prod_{j=1}^{p-1} \p{2k_j+j-n+2\mu S_z},
  \label{eq:E_mn_p_semifinal}
where $\Z_q$ denotes the set of $q$ smallest non-negative integers,
e.g.~$\Z_{\ell+1}=\set{0,1,\cdots,\ell}$.  We now expand
  \prod_{j=1}^{p-1} \p{2k_j+j-n+2\mu S_z}
  = \sum_{q=0}^{p-1} \p{2\mu S_z}^q
  \sum_{c\in\C_{p-1-q}\p{\Z_{p-1}}} \prod_{j\in c} \p{2k_{j+1}+j+1-n},
  \label{eq:product_expansion}
where $\C\p{k,\S}$ denotes the set of $k$-combinations (i.e.~subsets
with $k$ elements) of $\S$; the number of elements in
$\C_{p-1-q}\p{\Z_{p-1}}$ is
${ p-1 \choose p-1-q } = { p-1 \choose q }$.  Substituting
\eqref{eq:D_mn_k} and \eqref{eq:product_expansion} into
\eqref{eq:E_mn_p_semifinal}, we get that for $p>1$,
  = n^{\underline{p}} S_\bmu^{n-p-1} S_\mu^{m-p-1}
  \sum_{s=0}^2 f_{mnp\ell}^{(s)} \p{2\mu S_z}^s
  \sum_{q=0}^{p-1} g_{np\ell q} \p{2\mu S_z}^q,

@@

where
  f_{mnp\ell}^{(0)}
  &= \p{m-p-\ell} \p{n-p} \p{\ell+m-n} \p{2\ell-n+p}, \\
  f_{mnp\ell}^{(1)}
  &= \p{m-p-\ell} \p{n-p} \p{3\ell+m-2n+p}, \\
  f_{mnp\ell}^{(2)}
  &= \p{m-p-\ell} \p{n-p}, \\
  g_{np\ell q}
  &=
  \sum_{c\in\C_{p-1-q}\p{\Z_{p-1}}}
  \sum_{\substack{\v k\in\Z_{\ell+1}^{p-1}\\k_1\le k_2\le\cdots\le k_{p-1}}}
  \prod_{j\in c} \p{2k_{j+1}+j+1-n},
and for $p=1$ we can resolve \eqref{eq:E_mn_1_first} to get
  = \f16 mn \p{m-1} \p{3n-m-1} S_\bmu^{n-1} S_\mu ^{m-1} \p{m+2\mu S_z}.
  \label{eq:E_mn_1}
With reference to \eqref{eq:E_mn_p}-\eqref{eq:E_mn_1}, we finally have
  = mn S_\bmu^{n-1} S_\mu^{m-1}
  \sum_{q\in\set{0,1}} \p{m-n}^{1-q}\p{2\mu S_z}^q
  + \sum_{p=1}^{\min\set{m,n}-1} E_{\mu mn}^{(p)}.

>>>>

We now consider again squeezing the initial state $\ket{X}$, but this
time with the transverse field Hamiltonian
\begin{align}
  H_{\t{TVF}}
  \equiv \chi S_z^2- \Omega S_x.
\end{align}
To determine the coherent evolution of the operator
$S_\mu^m S_\nu^n S_z^n$, we now need to compute
\begin{align}
  \sp{H_{\t{TVF}}, S_\mu^m S_\nu^n S_z^n}_-
  = \chi \sp{S_z^2, S_\mu^m S_\nu^n S_z^n}_-
  - \Omega \sp{S_x, S_\mu^m S_\nu^n S_z^n}_-.
\end{align}
Here first commutator is identical to that for OAT, while

@@

  \label{eq:C_mn_E}

@@

@@

  = \f{n!}{\p{n-p}!}

@@

  D_{\mu,m-p,n-p}^{(k_p)} \prod_{j=1}^p \p{2k_j+j-n+2\mu S_z}.

@@

In principle, the expressions in \eqref{eq:D_mn} and
\eqref{eq:C_mn_E}-\eqref{eq:E_mn_p} provide a well-defined expansion
for the commutator $\sp{S_\mu^m,S_\nu^n}_- = C_{\mu mn}$, but the
expression for $E_{\mu mn}^{(p)}$ in \eqref{eq:E_mn_p} is -- put it
lightly -- quite a mess, involving a sum over $p$ mutually related
intermediate variables!  We therefore devote the rest of this section
to simplifying $E_{\mu mn}^{(p)}$.

Observing that we always have $k_1\le k_2\le\cdots\le k_p$, we can
reverse the order of the sums in \eqref{eq:E_mn_p} and re-index
$k_p\to\ell$ to get
\begin{align}
  E_{\mu mn}^{(p)}
  = \f{n!}{\p{n-p}!}
  \sum_{\ell=0}^{m-p-1} D_{\mu,m-p,n-p}^{(\ell)} \p{2\ell+p-n+2\mu S_z}
  \sum_{\substack{\v k\in\Z_{\ell+1}^{p-1}\\k_{p-1}\ge\cdots\ge k_1}}
  \prod_{j=1}^{p-1} \p{2k_j+j-n+2\mu S_z},
\end{align}
where $\Z_q$ denotes the set of $q$ smallest non-negative integers,
e.g.~$\Z_{\ell+1}=\set{0,1,\cdots,\ell}$.  Defining for shorthand
\begin{align}
  \sum_{\p{\v k,p,\ell}} X \equiv
  \sum_{\substack{\v k\in\Z_{\ell+1}^{p-1}\\k_{p-1}\ge\cdots\ge k_1}} X,
  &&
  F_{\mu np} \equiv p - n + 2\mu S_z,
\end{align}
we can say
\begin{align}
  E_{\mu mn}^{(p)}
  = \f{n!}{\p{n-p}!}
  \sum_{\ell=0}^{m-p-1} D_{\mu,m-p,n-p}^{(\ell)} \p{2\ell+F_{\mu np}}
  \sum_{\p{\v k,p,\ell}} \prod_{j=1}^{p-1} \p{2k_{p-j}-j+F_{\mu np}}.
  \label{eq:E_mn_p_k_sum}
\end{align}
We now define
\begin{align}
  f_{\mu np\ell}\p{k,q} \equiv \p{\ell-k+q} \p{\ell+k-q+F_{\mu np}},
\end{align}
and evaluate sums successively over $k_{p-1},k_{p-2},\cdots,k_2,k_1$
respectively from $k_{p-2},k_{p-3},\cdots,k_1,0$ to $\ell$, finding
\begin{align}
  \sum_{\p{\v k,p,\ell}} \prod_{j=1}^{p-1} \p{2k_{p-j}-j+F_{\mu np}}
  &= \sum_{\p{\v k;p-1;\ell}}
  \prod_{j=2}^{p-1} \p{2k_{p-j}-j+F_{\mu np}}
  f_{\mu n\ell}\p{k_{p-2},1} \\
  &= \f12 \sum_{\p{\v k,p-2,\ell}}
  \prod_{j=3}^{p-1} \p{2k_{p-j}-j+F_{\mu np}}
  f_{\mu np\ell}\p{k_{p-3},1} f_{\mu np\ell}\p{k_{p-3},2} \\
  &= \f1{r!} \sum_{\p{\v k,p-r,\ell}}
  \prod_{j=1+r}^{p-1} \p{2k_{p-j}-j+F_{\mu np}}
  \prod_{q=1}^r f_{\mu np\ell}\p{k_{p-1-r},q} \\
  &= \f1{\p{p-1}!} \prod_{q=1}^{p-1} f_{\mu np\ell}\p{0,q} \\
  &= { \ell + p - 1 \choose p - 1 }
  \prod_{q=1}^{p-1} \p{\ell-q+F_{\mu np}}.
\end{align}
Substitution of this result together with $D_{\mu,m-p,n-p}^{(\ell)}$
as given in using \eqref{eq:D_mn_k} into $E_{\mu mn}^{(p)}$ as
expressed in \eqref{eq:E_mn_p_k_sum} then gives us
\begin{align}
  E_{\mu mn}^{(p)}
  = \f{n!}{\p{n-p-1}!} S_\bmu^{n-p-1} S_\mu^{m-p-1} G_{\mu mnp}
  \label{eq:E_mn_p_G}
\end{align}
with
\begin{align}
  G_{\mu mnp}
  \equiv
  \sum_{\ell=0}^{m-p-1} { \ell + p - 1 \choose p - 1 }
  \p{m-p-\ell} \p{\ell+m-p+F_{\mu np}}
  \p{2\ell + F_{\mu np}}
  \prod_{q=1}^{p-1} \p{\ell-q+F_{\mu np}}.
\end{align}

% \vspace{3cm}

% With reference to \eqref{eq:E_mn_p_G}, we finally have
% \begin{align}
%   \sp{S_\mu^m, S_\bmu^n}_-
%   = C_{\mu mn}
%   = mn S_\bmu^{n-1} S_\mu^{m-1}
%   \sum_{q\in\set{0,1}} \p{m-n}^{1-q}\p{2\mu S_z}^q
%   + \sum_{p=1}^{\min\set{m,n}-1} E_{\mu mn}^{(p)}.
% \end{align}

% \begin{align}
%   \prod_{j=1}^{p-1} \p{2k_j+j-n+2\mu S_z}
%   = \sum_{q=0}^{p-1} \p{2\mu S_z}^q
%   \sum_{c\in\C_{p-1-q}\p{\Z_{p-1}}} \prod_{j\in c} \p{2k_{j+1}+j+1-n},
%   \label{eq:product_expansion}
% \end{align}
% where $\C\p{k,\S}$ denotes the set of $k$-combinations (i.e.~subsets
% with $k$ elements) of $\S$; the number of elements in
% $\C_{p-1-q}\p{\Z_{p-1}}$ is
% ${ p-1 \choose p-1-q } = { p-1 \choose q }$.
